Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I

### 2025-05-12 manuscript/manuscript.Rmd
Pivoting results section to be xamend first

maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
               "Mycosporine-2-glycine?")
nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
                  "Deoxyadenosine", "Deoxyguanosine?")
betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
                   "Threonine betaine?", "Proline betaine",
                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
                   "Betonicine", "Acetylcholine", "Phosphocholine")
non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
                 "2-O-alpha-D-Glucosylglycerol")
aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
              "Leucine/Isoleucine", "Valine", "Arginine")
bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
aa_ncs <- all_stans %>%
  filter(compound_type=="Amino Acid") %>%
  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
  distinct(compound_name, n_C, n_N) %>%
  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)

samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)
log_coef_df <- labeled_fracs %>%
  filter(!compound_name%in%maa_names) %>%
  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  select(compound_name, amendment, depth, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(half_max=-intercept/slope, scale=1/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))

splividis <- function(n, ...){
  vir <- viridis::viridis(n, ...)
  if(n%%2==0){
    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
  } else {
    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
    vir[head(index_vec, -1)]
  }
}
```

@@

## Environmental context of PARAGON cruise

PARAGON synthesis paper isn't out yet and I'm not sure it will be anytime soon

Particulate carbon and nitrogen numbers for PARAGON I (from Matt(?))

![](figures/borrowed/pcpn_PIandPII.png)

Data from Rhea showing ammonia/nitrite concentrations - I sampled at 25m and 175m (nitrite max, below ammonia max). Does this mean I was getting a lot of AOA users?

![](figures/borrowed/rhea_ammonia_concs.png)

Angel's IFCB data showing Hemiaulus disappearing over the course of the cruise (most of the Bacillariophyceae are Hemiaulus and she shows that in another plot)

![](figures/borrowed/angel_hemiaulus.png)

Fuyan reports the abundances of various bacterial groups (with the caveat that the SASVs have to have been found in the 4000m depth traps?)

![](figures/borrowed/fuyan_composition_bact.png)

## Universal bioavailability of ammonia at surface

The most rapid and extensive labeling was obtained during the ammonia-amended incubations of surface seawater. We observed significant labeling of glutamine, [and others] even at the T0 timepoint, indicating turnover times shorter than the time required to filter and flash-freeze the samples in liquid nitrogen (~20 minutes).

DISCUSSION: the speed at which ammonia labeling happens hints at a high flux, low concentration ammonia cycle in the surface ocean since uptake was too fast and too widespread to be a transcriptional response, indicating that they have ammonia transporters plus GS/GOGAT/GDH always active.

Synthesis rates, especially for betaines?

Aspartate labeled faster than glutamate - discuss pool sizes

Slow labeling of osmolytes - discuss pool sizes

```{r fig small heatmap compound groups}
class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", "Other nucleic acids", "MAAs", "Betaines", "All others")

labeled_fracs %>%
  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment)) %>%
  filter(amendment=="Amm") %>%
  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
  mutate(compound_class=case_when(
    compound_name%in%class_levels ~ compound_name,
    compound_name%in%maa_names~"MAAs",
    compound_name%in%aa_names~"Other amino acids",
    compound_name%in%nucleo_names~"Other nucleic acids",
    compound_name%in%betaine_names~"Betaines",
    TRUE~"All others"
  )) %>%
  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint_num, startime, depth)) %>%
  mutate(color_group=paste(depth, startime)) %>%
  ggplot(aes(x=timepoint_num, y=avg_frac, color=color_group, fill=color_group)) +
  geom_smooth(method = "glm", formula = y~x, method.args = list(family = "quasibinomial"),
              fill="grey80") +
  geom_point(pch=21, color="black") +
  facet_wrap(~compound_class, ncol=5) +
  scale_y_continuous(labels = scales::label_percent(), breaks = c(0, 0.5, 1)) +
  scale_x_continuous(breaks = c(0.3, 1, 3, 10, 26), labels = c(0.3, 1, 3, 10, 26)) +
  scale_color_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
  scale_fill_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
  labs(x="Hours after 15N added", y="Fraction of metabolite labeled", color=NULL) +
  theme_bw() +
  guides(color = guide_legend(nrow = 2, override.aes = list(linewidth=2)),
         fill = "none") +
  theme(legend.position = "inside",
        legend.position.inside = c(0.8, 1/6),
        legend.location = "plot")

labeled_fracs %>%
  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
  filter(amendment=="Amm") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("Surface seawater (25 meters)", "Seawater from below DCM (175m)"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
  mutate(compound_class=case_when(
    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine") ~ compound_name,
    compound_name%in%maa_names~"MAAs",
    compound_name%in%aa_names~"Other amino acids",
    compound_name%in%nucleo_names~"Other nucleic acids",
    compound_name%in%betaine_names~"Betaines",
    TRUE~"All others"
  )) %>%
  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint, startime, depth, tripl)) %>%
  ggplot() +
  geom_raster(aes(x=timepoint, y=tripl, fill=avg_frac)) +
  geom_vline(xintercept = c(1.5:4.5), color="grey80") +
  facet_nested(compound_class~depth+startime, scales = "free_y", switch="y") +
  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
                       limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1),
                       labels=c("Fully unlabeled", "", "50% labeled", "", "Fully labeled")) +
  scale_x_discrete(expand = expansion()) +
  scale_y_discrete(expand = expansion()) +
  labs(x=NULL, y=NULL, fill=NULL) +
  theme_bw() +
  theme(panel.spacing.x=unit(0.5, "lines"), 
        panel.spacing.y = unit(0, "lines"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y.left = element_text(angle=0, hjust = 1),
        strip.background.y = element_rect(fill=NA, color=NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "top", 
        legend.key.width = unit(1, "null")) +
  guides(
    fill=guide_colorbar(theme=theme(
      legend.frame = element_rect(color = "black"),
      legend.ticks = element_line(color="black"),
      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
    )
  )

ggsave(filename = "grouped_heatmap.png", device = "png", path = "manuscript/figures",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

![](figures/grouped_heatmap.png)

*Figure XX: Fraction of each metabolite or metabolite group containing at least one 15N label during the treatments amended with 15N ammonia after the number of hours shown on the x-axis. Darker colors indicate a larger percentage of the compound signal was isotopically labeled. Separate incubations were started at 6AM and 6PM with water from both 25 meters and 175 meters near Station ALOHA. Three biological replicates are stacked vertically at each timepoint.*

Same basic pattern in the evening except MAAs, guanosine, proline?

```{r fig MAA labeling patterns}
chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
                 "Also Porphyra-334?")
labeled_fracs %>%
  filter(compound_name%in%chosen_maas) %>%
  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
  arrange(compound_name) %>%
  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  filter(samp_type=="Smp") %>%
  filter(IS_type=="noIS") %>%
  filter(amendment%in%c("Amm", "Nit")) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
  ggplot() +
  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
              method.args=list(start=list(a=1, b=2, d=1),
                               control=nls.lm.control(maxiter = 500))) +
  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
  #             color="grey10", se=FALSE) +
  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
  #            fill="white", size=3) +
  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
             color="black", size=2, alpha=0.7) +
  facet_nested(compound_name~amendment) +
  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "")) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
                    aesthetics = c("color", "fill")) +
  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
  theme_bw() +
  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
        axis.title.x = element_blank(),
        legend.position = "inside",
        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
        legend.position.inside = c(1, 1),
        legend.justification = c(1, 1), 
        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))

ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
       width = 3, height = 6, units = "in", dpi = 300)
```

>>>>

maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
               "Mycosporine-2-glycine?")
nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
                  "Deoxyadenosine", "Deoxyguanosine?")
betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
                   "Threonine betaine?", "Proline betaine",
                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
                   "Betonicine", "Acetylcholine", "Phosphocholine")
non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
                 "2-O-alpha-D-Glucosylglycerol")
aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
              "Leucine/Isoleucine", "Valine", "Arginine")
bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Arginine", 
                  "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", 
                  "Guanosine", "Other nucleic acids", "MAAs", "Betaines", "All others")
aa_ncs <- all_stans %>%
  filter(compound_type=="Amino Acid") %>%
  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
  distinct(compound_name, n_C, n_N) %>%
  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)

samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)
```

@@

## Differences in nitrogen labeling at the surface between amendments

```{r}
xamend_metab_gp <- labeled_fracs %>%
  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
  filter(depth=="Surf") %>%
  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
  mutate(compound_class=case_when(
    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
    compound_name%in%maa_names~"MAAs",
    compound_name%in%aa_names~"Other amino acids",
    compound_name%in%nucleo_names~"Other nucleic acids",
    compound_name%in%betaine_names~"Betaines",
    TRUE~"All others"
  )) %>%
  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, amendment, timepoint, tripl)) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"),
                          labels=c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
  ggplot() +
  geom_raster(aes(x=timepoint, y=tripl, fill=avg_frac)) +
  geom_vline(xintercept = c(1.5:4.5), color="grey80") +
  facet_nested(compound_class~amendment, scales = "free_x", switch="y") +
  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
                       limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1),
                       labels=c("Fully unlabeled", "", "50% labeled", "", "Fully labeled")) +
  scale_x_discrete(expand = expansion()) +
  scale_y_discrete(expand = expansion()) +
  labs(x=NULL, y=NULL, fill=NULL) +
  theme_bw() +
  theme(panel.spacing.x=unit(0.5, "lines"), 
        panel.spacing.y = unit(0, "lines"),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y.left = element_text(angle=0, hjust = 1),
        strip.background.y = element_rect(fill=NA, color=NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "top", 
        legend.key.width = unit(1, "null")) +
  guides(
    fill=guide_colorbar(theme=theme(
      legend.frame = element_rect(color = "black"),
      legend.ticks = element_line(color="black"),
      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
    )
  )
ggsave("xamend_thaa_gp.png",plot = xamend_metab_gp, device = "png", path = "figures",
       width = 6.5, height = 5.5, units = "in", dpi = 300)
```

![](figures/xamend_thaa_gp.png)

*Figure 1: Fraction of each metabolite or metabolite group containing at least one 15N label during the surface seawater amended various substrates after the number of hours shown on the x-axis. Darker colors indicate a larger percentage of the compound signal was isotopically labeled. Three biological replicates are stacked vertically at each timepoint and the categories with multiple entries have been averaged. MAA = mycosporine-like amino acids, GMP = guanosine monophosphate.*

  - Ammonia
    - Really rapid labeling of glutamine/glutamate that aligns nicely with prior work
      - Aspartate labeled faster than glutamate likely due to small aspartate pool size
    - Surprisingly rapid labeling of citrulline that isn't reflected in arginine
      - Pool size problem (arginine >> citrulline)?
      - Indicates that urea cycle is only half-active?
    - Putative annotation of theanine based on labeling patterns
    - Rapid and heavy guanine labeling
      - Potentially being used as a storage compound
    - 10-40% of the N in the THAA (protein) pool ends up labeled by T26 (below)
      - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
  - Nitrate
    - Very little labeling, even in glutamate/glutamine
      - Surface community not "expecting" nitrate?
      - Redox problems with surface energy?
    - Guanine and MAA labeling exceeds glutamate labeling
      - Clear sign of a community *subset* using nitrate
        - Specifically the MAA synthesizers (euks) and those storing N in guanine (diatoms/haptos/dinos)
    - More extensive labeling (10-50% of the N) in the THAAs
      - Must be a really high flux through a small AA pool subset, e.g. euks specifically
  - Arginine
    - Clear uptake at T0 (20 minutes), 80-90% of the pool taken up at T10
      - Labeling signal decreases after 24 hours
    - Label obviously making it into citrulline so urea cycle is active
      - Don't really have good enough urea data to see labeling change there?
    - Big jump in labeling of other metabolites at T26 not seen at T10
      - No hypotheses why this is happening but it occurs in both morn/eve so not light signal
    - Maybe less storage in guanine specifically, labeling is on par with other NAs unlike ammonia
    - THAAs show clear usage of arginine in proteins
      - 30-50% arg in THAAs labeled @73hrs
      - Clear reuse of arginine backbone as glutamate (fully labeled glutamate) and maybe proline?
  - Guanosine monophosphate
    - Unable to measure GMP well enough directly?
    - Guanosine labeling visible at T3, definitely T10, maybe some decrease at T73 like Arg?
      - Guanosine pool never gets super heavily labeled
    - Has a VERY similar pattern to arginine
      - Also a big jump at T26 with basically no labeling at T10
      - Implies mostly converted into ammonia as well and used that way
    - Reuse of carbon backbone for THAAs though
      - Histidine 13C5, 15N3
      - Tyrosine 13C4, 15N1
      - Phenylalanine 13C4, 15N1

```{r thaa data}
thaa_quant <- read_csv("quant_data.csv")
metathaa <- read_csv("metathaa.csv") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve")))

all_thaa_plot <- thaa_quant %>%
  filter(!str_detect(iso_name, "2H")) %>%
  left_join(metathaa) %>%
  filter(samp_type=="Smp") %>%
  left_join(aa_ncs) %>%
  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
  mutate(n_N_unlab=n_N-n_N_lab) %>%
  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
  left_join(metathaa) %>%
  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
                          labels=c("Initial", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
  summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
  mutate(name=factor(name, levels=c("nM_N_unlab", "nM_N_lab"), labels = c("Unlabeled", "Labeled"))) %>%
  ggplot(aes(x=amendment, y=total_uM, fill=name)) +
  geom_col(color="black", width=1) +
  theme_bw() +
  scale_x_discrete(expand = expansion()) +
  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
  scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)

plot_iso_levels <- c(
  "Unlabeled", 
  "<sup>15</sup>N<sub>1</sub>", 
  "<sup>15</sup>N<sub>2</sub>", 
  "<sup>15</sup>N<sub>3</sub>", 
  "Entirely <sup>15</sup>N",
  "<sup>13</sup>C<sub>1</sub>", 
  "<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
  "<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>", 
  "<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
  "Entirely <sup>13</sup>C<br>and <sup>15</sup>N", 
  "Other labels")
plot_iso_colors <- c(
  "grey80",
  "#e0ecf6",
  "#97bfe0",
  "#377EB8",
  "#285b84",
  "#ffb4b5",
  "#d7b6dc",
  "#FFFF33",
  "#4DAF4A",
  "#984EA3",
  "grey20"
)
# "#E41A1C" "#377EB8" "#4DAF4A" "#984EA3" "#FF7F00" "#FFFF33" "#A65628" "#F781BF" "#999999"
thaa_plot_data <- thaa_quant %>%
  left_join(metathaa) %>%
  filter(samp_type=="Smp") %>%
  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint)) %>%
  filter(compound_name%in%c("Glutamic acid", "Arginine", "Histidine", "Phenylalanine")) %>%
  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glutamine + Glutamate", compound_name)) %>%
  mutate(plot_iso_name=case_when(
    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
    str_detect(iso_name, "13C0, 15N1")~"<sup>15</sup>N<sub>1</sub>",
    str_detect(iso_name, "13C0, 15N2")~"<sup>15</sup>N<sub>2</sub>",
    str_detect(iso_name, "Histidine, 13C0, 15N3")~"Entirely <sup>15</sup>N",
    str_detect(iso_name, "13C0, 15N3")~"<sup>15</sup>N<sub>3</sub>",
    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Entirely <sup>15</sup>N",
    str_detect(iso_name, "13C1, 15N0")~"<sup>13</sup>C<sub>1</sub>",
    str_detect(iso_name, "13C1, 15N1")~"<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Entirely <sup>13</sup>C and <sup>15</sup>N",
    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
    iso_name=="Phenylalanine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",
    iso_name=="Histidine, 13C5, 15N3"~"<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
    TRUE~"Other labels"
  )) %>%
  mutate(plot_iso_name=factor(plot_iso_name, levels=plot_iso_levels)) %>%
  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP")))
arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
  geom_col(aes(x=amendment, y=avg_env_conc_nm, fill=plot_iso_name, label=iso_name),
           position="fill",
           color="black", width=1) +
  facet_wrap(~compound_name) +
  theme_bw() +
  scale_x_discrete(expand = expansion()) +
  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
                     expand = expansion()) +
  scale_fill_manual(breaks = plot_iso_levels, values = plot_iso_colors) +
  labs(x=NULL, y="Fraction of THAA labeled", fill=NULL) +
  theme(
    # axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5),
    legend.text = element_markdown()
  )
total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.3, 0.7))
total_thaa_gp
ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "figures",
       width = 6.5, height = 5, units = "in", dpi = 300)
```

*Figure 2: Stacked bar plots showing the fraction of various total hydrolyzable amino acids (THAAs) labeled at the end of the incubations across the four amendments. Total uM nitrogen (N) in the protein pool is shown in the top plot while the fraction of individual amino acids labeled is shown in the rows below. Colors denote specific labeling patterns detected in the THAAs.*

## Minimal diel signal observed except in the UV-reactive mycosporine-like amino acids
